A coherent and comprehensive dating system for stars in the Milky Way

### 2017-02-06 documents/ms.tex
adding table of data sets

% Missions
\newcommand{\Kepler}{{\it Kepler}}
\newcommand{\kepler}{\Kepler}
\newcommand{\corot}{{\it CoRoT}}
\newcommand{\Ktwo}{{\it K2}}
\newcommand{\ktwo}{\Ktwo}
\newcommand{\TESS}{{\it TESS}}
\newcommand{\tess}{{\it TESS}}
\newcommand{\LSST}{{\it LSST}}
\newcommand{\lsst}{{\it LSST}}
\newcommand{\Wfirst}{{\it Wfirst}}
\newcommand{\wfirst}{{\it Wfirst}}
\newcommand{\SDSS}{{\it SDSS}}
\newcommand{\sdss}{{\it SDSS}}
\newcommand{\PLATO}{{\it PLATO}}
\newcommand{\plato}{{\it PLATO}}
\newcommand{\TGAS}{{\it TGAS}}
\newcommand{\tgas}{{\it TGAS}}
\newcommand{\APOGEE}{{\it APOGEE}}
\newcommand{\apogee}{{\it APOGEE}}
\newcommand{\gaia}{{\it Gaia}}
\newcommand{\Gaia}{{\it Gaia}}

@@

% Paper-specific
\newcommand{\nkictags}{2000}

@@

\author{%
    Ruth Angus\altaffilmark{\columbia,}\altaffilmark{\simons},
   David Hogg\altaffilmark{\nyu,}\altaffilmark{\mpia,}\altaffilmark{\cca},
    \etal
   % Melissa Ness\altaffilmark{\mpia},
   % David Kipping\altaffilmark{\columbia}, \etal
}

@@

\begin{landscape}
\begin{table}
\begin{center}
\caption{A list of all dating methods with the types of stars they apply to
    and their effective or physical natures.}
\begin{tabular}{lcccc}
    Method & Observable & Underlying cause & Applicable to... & Effective or
    Physical model? \\
    \hline
    Rotation & $P_{\mathrm{rot}}$ & Magnetic activity & $<$ 4 Gyr, MS &
    Effective \\
    Activity & Radio, H$\alpha$ \& X-ray & Magnetic activity & $<$
    5 Gyr, low mass & Effective \\
    Time variability & Time variability & ? & ? & ? \\
    Isochrones & $T_{\mathrm{eff}}$, $M_v$ & Core/shell fusion &
    Subgiants/giants & Physical (model dependent) \\
    Asteroseismology & Frequencies & Core/shell fusion & Old, low
    mass, Giants & Physical (model dependent) \\
    White dwarf cooling & Cooling & $M_v$ & WDs (\& companions) &
    Physical (1 free parameter) \\
    $J_z$, $J_{\phi}$, $J_r$ & Position \& velocity & Dynamic heating & Disk
    stars & Physical (1 free parameter) \\
    Galactocentric position & Position & Formation history &
    All stars & Effective \\
    Coeval pairs & Position \& velocity & Disruption & $<$ 1 Gyr, disk &
    0 free parameters \\
    Open Clusters & Disruption & Positions \& velocities & $<$ 2 Gyr &
    0 free parameters \\
    Exoplanet dynamics & Stability \& eccentricity & Interactions & $<$ 500
    Myrs & 0 free parameters \\
    Universe age & All of the above & Finite age of the Universe &
    All stars & 0 free parameters \\
    C/N ratio & C/N ratio & Convective processes & Red giants, A \& F &
    Effective \\
    Chemical abundances & Abundances & ISM enrichment & $>$ 8 Gyr?
    & Effective \\
    Lithium abundance & L abundance & $L$ depletion & $<$ 500
    Myrs? & Effective \\
    Nuclear isotopes & ? & ? & ? & ? \\
    Unusual HRD objects & $T_{\mathrm{eff}}$, $M_v$ & Mergers, ? &
    Special cases & ? \\
\end{tabular}
\end{center}
\end{table}
\label{tab:dating_methods}
\end{landscape}

@@

\section{Data}
\label{sec:data}
We assembled a sample of stars that have at least two of the four dating
methods available.
Our main asteroseismic sample targets comprise the $~$500 short cadence
\kepler\ prime targets.
These stars all have light curves and therefore rotation periods (or potential
rotation periods).
The majority are also in Gaia DR1, therefore have parallaxes, positions and
2-dimensional velocities, which can provide constraints on their vertical
actions.
A subset of these stars have spectroscopic information and therefore make
excellent isochronal targets.
Our main rotation sample is the \nkictgas\ stars in both the \kepler\ input
catalogue (KIC) and \TGAS.
These stars have potential rotation periods, parallaxes, positions and
2-dimensional velocities.
Our main dynamics data set is the intersection of the \apogee\ and \tgas\
catalogues.
The majority of these stars have ages from {\tt the cannon} \cite{Ness2015},
plus \gaia\ parallaxes, positions and proper motions as well as radial
velocities, providing vertical actions calculated from the full 3-dimensional
phase-space.
They are mostly red giants.

>>>>

% Missions
\newcommand{\Kepler}{{\it Kepler}}
\newcommand{\kepler}{\Kepler}
\newcommand{\kic}{\KIC}
\newcommand{\KIC}{\KIC}
\newcommand{\corot}{{\it CoRoT}}
\newcommand{\Ktwo}{{\it K2}}
\newcommand{\ktwo}{\Ktwo}
\newcommand{\TESS}{{\it TESS}}
\newcommand{\tess}{{\it TESS}}
\newcommand{\LSST}{{\it LSST}}
\newcommand{\lsst}{{\it LSST}}
\newcommand{\Wfirst}{{\it Wfirst}}
\newcommand{\wfirst}{{\it Wfirst}}
\newcommand{\SDSS}{{\it SDSS}}
\newcommand{\sdss}{{\it SDSS}}
\newcommand{\PLATO}{{\it PLATO}}
\newcommand{\plato}{{\it PLATO}}
\newcommand{\TGAS}{{\it TGAS}}
\newcommand{\tgas}{{\it TGAS}}
\newcommand{\APOGEE}{{\it APOGEE}}
\newcommand{\apogee}{{\it APOGEE}}
\newcommand{\gaia}{{\it Gaia}}
\newcommand{\Gaia}{{\it Gaia}}
\newcommand{\rave}{{\it RAVE}}
\newcommand{\Rave}{{\it RAVE}}

@@

% Paper-specific
\newcommand{\nkictgas}{2000}

@@

\author{%
    Ruth Angus\altaffilmark{\columbia,}\altaffilmark{\simons},
   David Hogg\altaffilmark{\nyu,}\altaffilmark{\cca,}\altaffilmark{\mpia},
    \etal
   % Melissa Ness\altaffilmark{\mpia},
   % David Kipping\altaffilmark{\columbia}, \etal
}

@@

\begin{landscape}
\begin{table}
\begin{center}
\caption{A list of all dating methods with the types of stars they apply to
    and their effective or physical natures.}
\begin{tabular}{lcccc}
    Method & Observable & Underlying cause & Applicable to... & Effective or
    Physical model? \\
    \hline
    Rotation & $P_{\mathrm{rot}}$ & Magnetic activity & $<$ 4 Gyr, MS &
    Effective \\
    Activity & Radio, H$\alpha$ \& X-ray & Magnetic activity & $<$
    5 Gyr, low mass & Effective \\
    Time variability & Time variability & ? & ? & ? \\
    Isochrones & $T_{\mathrm{eff}}$, $M_v$ & Core/shell fusion &
    Subgiants/giants & Physical (model dependent) \\
    Asteroseismology & Frequencies & Core/shell fusion & Old, low
    mass, Giants & Physical (model dependent) \\
    White dwarf cooling & Cooling & $M_v$ & WDs (\& companions) &
    Physical (1 free parameter) \\
    $J_z$, $J_{\phi}$, $J_r$ & Position \& velocity & Dynamic heating & Disk
    stars & Physical (1 free parameter) \\
    Galactocentric position & Position & Formation history &
    All stars & Effective \\
    Coeval pairs & Position \& velocity & Disruption & $<$ 1 Gyr, disk &
    0 free parameters \\
    Open Clusters & Disruption & Positions \& velocities & $<$ 2 Gyr &
    0 free parameters \\
    Exoplanet dynamics & Stability \& eccentricity & Interactions & $<$ 500
    Myrs & 0 free parameters \\
    Universe age & All of the above & Finite age of the Universe &
    All stars & 0 free parameters \\
    C/N ratio & C/N ratio & Convective processes & Red giants, A \& F &
    Effective \\
    Chemical abundances & Abundances & ISM enrichment & $>$ 8 Gyr?
    & Effective \\
    Lithium abundance & L abundance & $L$ depletion & $<$ 500
    Myrs? & Effective \\
    Nuclear isotopes & ? & ? & ? & ? \\
    Unusual HRD objects & $T_{\mathrm{eff}}$, $M_v$ & Mergers, ? &
    Special cases & ? \\
    Twins & ? & ? & ? & ? \\
\end{tabular}
\end{center}
\end{table}
\label{tab:dating_methods}
\end{landscape}

@@

\section{Data}
\label{sec:data}
We assembled a sample of stars that have at least two of the four dating
methods available.
Our main asteroseismic sample targets comprise the $~$500 short cadence
\kepler\ prime targets.
These stars all have light curves and therefore rotation periods (or potential
rotation periods).
The majority are also in Gaia DR1, therefore have parallaxes, positions and
2-dimensional velocities, which can provide constraints on their vertical
actions.
A subset of these stars have spectroscopic information and therefore make
excellent isochronal targets.
Our main rotation sample is the \nkictgas\ stars in both the \kepler\ input
catalogue (KIC) and \TGAS.
These stars have potential rotation periods, parallaxes, positions and
2-dimensional velocities.
Some of these stars, particularly those with confirmed planets, also have
spectroscopic information and could provide isochronal ages.
Our main isochronal data set is the intersection of the \apogee\ and \rave\
catalogues with \tgas.
Relatively precise isochronal ages can be inferred from the spectroscopic
stellar properties of these stars, the majority which have ages from {\it the
cannon} \citep{Ness2015, Casey2016}.
They also have \gaia\ parallaxes, positions and proper motions as well as
radial velocities, providing vertical actions calculated from the full
3-dimensional phase-space.

\begin{landscape}
\begin{table}
\begin{center}
\caption{A summary of the data sets used in this project}
\begin{tabular}{lccc}
    Sample & Catalogue intersection &  Methods available \\
    \hline
    Asteroseismic & \citet{Chaplin2014} \& TGAS & Asteroseismology, rotation,
    isochrones, dynamics \\
    Rotational & KIC \& TGAS & Rotation, isochrones, dynamics \\
    Isochronal & RAVE, APOGEE \& TGAS & Isochrones, dynamics \\
\end{tabular}
\end{center}
\end{table}
\label{tab:data}
\end{landscape}
